### 2016-11-07 methodsPaper/tex/materialsAndMethods.tex
Edits to Materials & Methods

As with the original ITHIM implementation we employed a comparative
risk analysis (CRA) across scenarios defined by mean active
transportation time, i.e., walking and cycling.  To do so we
investigate the proportional change in national disease burden between
the baseline model and a given alternative scenario, i.e., the
population attributable fraction, $\af$.  The population attributable
fraction is defined below and computed for each disease included in
the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
Diabetes and Cardiovascular disease.

\begin{equation}
\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \, \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
exposure and non-travel-related exposure.  Non-travel-related exposure
is modeled with a \logNormal{} distribution with the age-sex specific
mean ratios for non-travel-related exposure are estimated from the
ATUS \cite{ATUS}.  The referent group mean is treated as a variable
and the coefficient of variation is treated as constant across age-sex
strata and estimated with ATUS.  The distribution for travel-related
exposure is estimated from the time spent walking or cycling, i.e.,
active transport time, which as in the original ITHIM model has
a \logNormal{} distribution with constant coefficient of variation
across age-sex strata.  Travel-related exposure is estimated from
active transport time using assumptions about how much physical
activity is required for cycling and walking.

\paragraph{Approximation of $\af$ Integral}

\begin{equation}
1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
\end{equation}

\paragraph{Disease Burden Estimation}
\begin{equation}
B_1 = B_0(1-\af)\frac{\RR_1}{\sum_j \RR_j}
\end{equation}

$B_j = \gamma_jB_1$, where $\gamma_i = \frac{\RR_i}{\RR_1}$.  The
overall burden is then computed as the sum across quintiles,
$B=\sum B_j$.  Finally, $B_0$ is estimated from CDC data\ldots

@@

% does this involve using the CDC Wonder data? we could add a bit here
% about it, but if this is just national estimtes and shifts perhaps
% we don't need it.

>>>>

As with the original ITHIM implementation we employed a comparative
risk analysis (CRA) across scenarios defined by mean active
transportation time, i.e., walking and cycling.  To do so we
investigate the proportional change in national disease burden between
the baseline model and a given alternative scenario, i.e., the
population attributable fraction, $\af$.  The population attributable
fraction is defined below and computed for each disease included in
the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
Diabetes and Cardiovascular disease.  For readability we omit an index
indicating disease.

\begin{equation}
\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \,
  \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x} \label{paf}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
and non-travel-related physical activity.

\paragraph{Non-Travel-Related Physical Activity}

\begin{equation}
X_{ij}^{\mathrm{non-travel}} \sim \logNormalMath\left(\mu_{ij}, \gamma \mu_{ij}\right)
\end{equation}

Within each age-sex class the non-travel-related exposure is modeled
with a \logNormal{} distribution.  The age-sex specific mean ratios,
$r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
American Time USe Survey data \cite{ATUS}.

\paragraph{Travel-Related Physical Activity}

The distribution for travel-related exposure is estimated from the
time spent walking or cycling, i.e., active transport time, which as
in the original ITHIM model is assigned a \logNormal{} distribution
with constant coefficient of variation across age-sex classes.
Travel-related exposure is estimated from active transport time using
assumptions about how much physical activity is required for cycling
and walking.

\paragraph{Estimation of Disease Burden}

To estimate the disease burden we first approximate the population
attributable fraction through a discretization of the integral in
equation \ref{paf}.

With a simulated distribution for total physical activity in hand we
compute the empirical percentiles in the baseline and alternate
scenarios, $\mathbf{x}$ and $\mathbf{x}^\prime$. The proportional
change in disease burden, $\delta$, may be approximated as follows,
where $R$ is the risk function for total activity.

\begin{equation}
\delta = 1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
\end{equation}

The disease burden, in this case DALYs, is found for each of the
percentiles using equation \ref{burden}, using $B_0$, the overall nation-wide burden for
the baseline.  The overall burden is then computed as the sum across
quintiles, $B=\sum_{j=1}^n B_j$.

\begin{equation}
B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_k}\label{burden}
\end{equation}

@@
